Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-03 ms/SI.Rmd
Don't fiddle with CRS
Since CRS is same for all data when loaded (JGD2000)

# Set CRS
japan_crs <- sf::st_crs(japan)

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set CRS
  sf::st_set_crs(japan_crs) %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )

biodiv_ferns_endemic_spatial <-
  biodiv_ferns_endemic_spatial %>% 
  # Set CRS
  sf::st_set_crs(japan_crs) %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super"))
  )

@@

```{r protected-map, fig.width = 7, fig.height = 9}
# Define function to make maps showing overlap of protected areas and grid-cells with high biodiversity
# These depend on various objects being in the current environment!
# pass the dots to filter data
make_protected_plot <- function(part, variable, ...) {
  ggplot() +
    geom_sf(data = japan, lwd = 0.10, fill = "white") +
    geom_sf(data = protected_areas %>% mutate(status = str_to_sentence(status)), aes(fill = status), color = "transparent") +
    geom_sf(
      data = biodiv_ferns_spatial %>% filter(...) %>% st_set_crs(japan_crs), 
      fill = "transparent", color = "black", lwd = 0.20) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
      hjust = "inward", vjust = "inward", color = "white") +
    scale_fill_manual(values = protection_cols, name = "Protection status") +
    scale_x_continuous(breaks = c(125, 135, 145)) +
    scale_y_continuous(breaks = c(25, 35, 45)) +
    map_theme()
}

@@

```{r deer-map, fig.width = 7, fig.height = 9}
# Define function to make maps showing overlap of areas with deer and grid-cells with high biodiversity
# These depend on various objects being in the current environment!
make_deer_plot <- function(part, variable, ...) {
  ggplot() +
    geom_sf(data = japan, lwd = 0.10, fill = "white") +
    geom_sf(data = deer_range %>% filter(range == "estimated") %>% mutate(range = str_to_sentence(range)), aes(fill = range), color = "transparent") +
    geom_sf(data = deer_range %>% filter(range == "2003"), aes(fill = range), color = "transparent") +
    geom_sf(data = deer_range %>% filter(range == "1978"), aes(fill = range), color = "transparent") +
    geom_sf(
      data = biodiv_ferns_spatial %>% filter(...) %>% st_set_crs(japan_crs), 
      fill = "transparent", color = "black", lwd = 0.20) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
      hjust = "inward", vjust = "inward", color = "white") +
    scale_fill_manual(values = deer_cols, name = "Range data source") +
    scale_x_continuous(breaks = c(125, 135, 145)) +
    scale_y_continuous(breaks = c(25, 35, 45)) +
    map_theme()
}

>>>>

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )

biodiv_ferns_endemic_spatial <-
  biodiv_ferns_endemic_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super"))
  )

@@

```{r protected-map, fig.width = 7, fig.height = 9}
# Define function to make maps showing overlap of protected areas and grid-cells with high biodiversity
# These depend on various objects being in the current environment!
# pass the dots to filter data
make_protected_plot <- function(part, variable, ...) {
  ggplot() +
    geom_sf(data = japan, lwd = 0.10, fill = "white") +
    geom_sf(data = protected_areas %>% mutate(status = str_to_sentence(status)), aes(fill = status), color = "transparent") +
    geom_sf(
      data = biodiv_ferns_spatial %>% filter(...), 
      fill = "transparent", color = "black", lwd = 0.20) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
      hjust = "inward", vjust = "inward", color = "white") +
    scale_fill_manual(values = protection_cols, name = "Protection status") +
    scale_x_continuous(breaks = c(125, 135, 145)) +
    scale_y_continuous(breaks = c(25, 35, 45)) +
    map_theme()
}

@@

```{r deer-map, fig.width = 7, fig.height = 9}
# Define function to make maps showing overlap of areas with deer and grid-cells with high biodiversity
# These depend on various objects being in the current environment!
make_deer_plot <- function(part, variable, ...) {
  ggplot() +
    geom_sf(data = japan, lwd = 0.10, fill = "white") +
    geom_sf(data = deer_range %>% filter(range == "estimated") %>% mutate(range = str_to_sentence(range)), aes(fill = range), color = "transparent") +
    geom_sf(data = deer_range %>% filter(range == "2003"), aes(fill = range), color = "transparent") +
    geom_sf(data = deer_range %>% filter(range == "1978"), aes(fill = range), color = "transparent") +
    geom_sf(
      data = biodiv_ferns_spatial %>% filter(...), 
      fill = "transparent", color = "black", lwd = 0.20) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
      hjust = "inward", vjust = "inward", color = "white") +
    scale_fill_manual(values = deer_cols, name = "Range data source") +
    scale_x_continuous(breaks = c(125, 135, 145)) +
    scale_y_continuous(breaks = c(25, 35, 45)) +
    map_theme()
}

### 2021-09-03 ms/manuscript.Rmd
Don't fiddle with CRS
Since CRS is same for all data when loaded (JGD2000)

We used a list of all non-hybrid fern specimens deposited at TNS (herbarium codes follow Thiers2021). 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].

@@

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
japan_crs <- sf::st_crs(japan)

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set CRS
  sf::st_set_crs(japan_crs) %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )
```

>>>>

We used a list of all non-hybrid fern specimens deposited at TNS (herbarium codes follow Thiers2021). 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
Points were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.
The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.

@@

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )
```
